Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2021-01-26 xrb2/paper.tex
minor changes

%======================================================================
% Abstract and Keywords
%======================================================================
\begin{abstract}
We continue to investigate two-dimensional laterally propagating flames in type I X-ray bursts using fully compressible hydrodynamics simulations.  In the current study we relax previous approximations where we artificially boosted the flames. We now use more physically realistic reaction rates, thermal conductivities, and rotation rates, exploring the effects of neutron star rotation rate and thermal structure on the flame.  We find that at lower rotation rates the flame becomes harder to ignite, whereas at higher rotation rates the nuclear burning is enhanced by increased confinement from the Coriolis force and the flame propagates steadily. At higher crustal temperatures, the flame moves more quickly and accelerates as it propagates through the atmosphere. If the temperature is too high, instead of a flame propagating across the surface the entire atmosphere burns steadily. All of the software used for these simulations is freely available.

@@

We investigate the effect of rotation rate on the flame. With the
exception of IGR J17480-2446 (\citealt{altamirano2010discovery}, spinning at $11~\mathrm{Hz}$), most observations of
XRBs come from sources with rotation rates of $200-600~\mathrm{Hz}$
\citep{bilous:2019,galloway:2020}. There are a number of
factors that could explain this lack of observations below $200~\mathrm{Hz}$. It
could be that there is some physical process which inhibits the flame
ignition and/or spread at lower rotation rates. It could be that burst
oscillations at lower rotation rates are smaller in amplitude and
therefore more difficult to detect. It could be that it does not have
anything to do with the flame at all, but that neutron stars in
accreting low mass X-ray binaries rarely have rotation rates below $200~\mathrm{Hz}$ 
\MarginPar{look into this?}.

@@

We use the Castro hydrodynamics code \citep{castro,castro_joss} and
the simulation framework introduced in \citet{flame_wave1}.  The
current simulations are all performed in a two-dimensional
axisymmetric geometry.  For these axisymmetric simulations, we add an
additional geometric source term from \citet{bernard-champmartin} that
captures the effects of the divergence of the flux operating on the
azimuthal unit vector.  This term is a small correction, but was
missing from our previous simulations.  The simulation framework
initializes a fuel layer in hydrostatic equilibrium, laterally
blending a hot model on the left side of the domain (the coordinate
origin) and a cool model on the right.  The initial temperature
gradient between the hot and cool fluids drives a laterally
propagating flame through the cool fuel.  In our original set of
calculations (in in \citet{flame_wave1}), in order to make the 
simulations computationally feasible we artificially boosted the flame speed by adjusting the
conductivity and reaction rate to produce a flame moving 5--10$\times$
faster than the nominal laminar flame speed.  We also used high
rotation rates to reduce the lateral lengthscale at which the Coriolis
force balances the lateral flame spreading in order to reduce the size 
of the simulation domain. The port of \castro\ to GPUs~\citep{sc20_gpu} 
significantly improved its overall performance, enabling
us to run these new simulations without the previous approximations while
continuing to resolve the burning front. For these
simulations, we no longer boost the flame speed---the true
conductivities (taken from \citealt{Timmes00}) and reaction rates are
used. We are also able to use slower, more physically realistic rotation rates.
  We continue to use a 13-isotope $\alpha$-chain to describe the
helium burning.

@@

\begin{figure}
	\plotone{maxima_vs_time.pdf}
	\caption{\label{fig:maxima}Estimates of temperature (left panel) and nuclear energy generation 
	rate (right panel) in the burning region as functions of time. The quantities on the vertical 
	axes are the mass-weighted averages defined in Equation~\ref{eqn:top_percentile_avg} for the temperature (left panel) and energy generation rate (right panel).}
\end{figure}

For some of our analysis, we would like to have a means of estimating the temperature ($T$) and 
nuclear energy generation rate ($\enucdot$) in the burning region of each simulation. For this 
purpose, we define the mass-weighted average $\repavg{Q}$ of some quantity $Q$ to be
\begin{equation}
	\label{eqn:top_percentile_avg}
	\repavg{Q} \equiv \frac{\sum_{c_i} m(c_i) Q(c_i)}{\sum_{c_i} m(c_i)}; ~c_i \in C_{99}(Q).
\end{equation}
\noindent Here, $C_{99}(Q)$ is the set of grid cells with $Q$ values in the top percentile, and 
$Q(c_i)$ is the value of $Q$ in cell $c_i$ (with $m(c_i)$ being the total mass in cell $c_i$). By 
using $\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire 
simulation domain allows us to track changes over the domain as a whole rather than at a single 
localized point. This will therefore be a better reflection of the overall behavior of the flame 
rather than of a single localized fluctuation. Figure \ref{fig:maxima} shows $\repavg{T}$ and 
$\repavg{\enucdot}$ as functions of time for the subset of our runs that achieve a propagating 
flame. This figure is referenced throughout the subsequent sections.

@@

We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms 
of temperature but differing rotation rates. We saw in \citet{flame_wave1} that 
increasing the neutron star rotation rate reduces the horizontal lengthscale of the flame. An 
estimate of this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius 
may be thought of as the scale over which the balance between the Coriolis force and horizontal 
pressure gradient becomes important, and is approximately given by
\begin{equation}
	\label{eqn:rossby}
	L_R \approx \frac{\sqrt{g H_0}}{\Omega},
\end{equation}
where $g$ is the gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
neutron star rotation rate. In Figure \ref{fig:time_series_enuc_500} and Figure 
\ref{fig:time_series_enuc_1000}, we use the $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates. Taking the edge at greatest radius of the bright teal/green region where the most significant energy generation is occurring as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be 
reduced compared to the lower rotation $500~\mathrm{Hz}$ run ($\tt{F500}$). From Equation 
\ref{eqn:rossby}, we can see that increasing the rotation rate from $500~\mathrm{Hz}$ to 
$1000~\mathrm{Hz}$ should decrease $L_R$ by a factor of two, and that the greater confinement from 
the Coriolis force should reduce the horizontal extent of the flame by a similar factor. However, 
the Rossby radius is only an approximate measure of this horizontal lengthscale, and in our 
simulations we see that this scaling does not work so well for all rotation rates. The simulations 
seem to follow the theoretical scaling described in Equation \ref{eqn:rossby} more closely at 
higher rotation rates ($1000~\mathrm{Hz}$ and higher), based on the results of \citet{flame_wave1}.

@@

The {\tt F500} and {\tt F1000} runs both qualitatively resemble the flame structure in 
\citet{flame_wave1} --- a laterally propagating flame that is lifted off of the bottom of the fuel 
layer --- but they differ in their burning structures. Figures~\ref{fig:time_series_500} and 
\ref{fig:time_series_1000} show time series of the mean molecular weight, $\bar{A}$, for the {\tt 
F500} and {\tt F1000} runs.  Compared to those in \cite{flame_wave1}, ashes behind the flame do not 
reach as high atomic weights. This is not surprising, since those previous runs artificially 
boosted the reaction rates.  Comparing these two new runs, the burning is much more evolved for the 
higher rotation rate, and the ash is actually able to move ahead of the flame front (visible in the 
Figure  \ref{fig:time_series_1000} $100~\mathrm{ms}$ snapshot). We believe that this is because the increased 
rotation better confines the initial perturbation and subsequent expansion from the burning, 
increasing the temperature and density in the flame front such that the reaction rate increases, 
which allows the reactions to progress further. The $\dot{e}_\mathrm{nuc}$ plots in Figure 
\ref{fig:time_series_enuc_1000} also support this interpretation, with the region of the 
flame front nearest to the crust in the {\tt F1000} run reaching higher $\dot{e}_\mathrm{nuc}$ values 
than for the {\tt F500} run in Figure~\ref{fig:time_series_enuc_500}. In contrast to {\tt F500} 
and {\tt F1000}, the lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition 
for {\tt F250} also aligns with the reasoning given above, with the lower rotation in this case 
potentially leading to insufficient confinement such that the temperature and density required for 
ignition is not achieved. In this scenario, another source of confinement (e.g.\ magnetic fields, see \citep{art-2016-cavecchi-etal}) 
would need to take over at lower rotation rates to allow a burst to occur, at least for the initial 
model used here. Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), it is also possible that we simply cannot confine the flame sufficiently with our current domain width. We see in Figure~\ref{fig:flame_speeds_1} that the {\tt F500} flame took longer to achieve steady propagation than the {\tt F1000} flame. It may therefore also be that we did not run our simulation for long enough to see the {\tt F250} flame achieve the conditions required for ignition and steady propagation.

Burning in the {\tt F500} and {\tt F1000} runs is concentrated in a dense region with circular 
motion. In Figure~\ref{fig:urho}, which compares the horizontal $r$-velocity $u$, density $\rho$ and the 
nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the {\tt F250}, {\tt F500}, and {\tt F1000} 
runs, most of the burning for each of the simulations occurs in a high density region $\rho > 3 
\times 10^5 {\rm g\,cm^{-3}}$. The fluid in this dense, high energy generation region undergoes 
vortical motion, shown in the Figure~\ref{fig:uv} phase plots comparing $u$, the vertical $z$-velocity $v$ and $\dot{e}_{\rm 
nuc}$. This most likely corresponds to the leading edge of the flame where fresh fuel is being entrained.
This feature is not developed in the $250~\mathrm{Hz}$ flame in Figure~\ref{fig:uv} (left panel); it could potentially develop 
at later times (past the point at which we terminated our simulation), or the flame could just 
fizzle out and fail to ignite entirely.

@@

Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures ($T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$) at $t = 70~\mathrm{ms}$. We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front. $\tt{F500\_3E8}$ (Figure~\ref{fig:compare_500Hz_abar}, bottom panel) --- the hottest run to form a clear burning front --- has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}). It also reaches slightly higher $\bar{A}$ values than the two cooler runs. The Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plots of $\tt{F500}$ (left) and $\tt{F500\_3E8}$ (middle) also reflect these $\bar{A}$ features, with $\tt{F500\_3E8}$ reaching slightly higher $\bar{A}$ values. $\tt{F500\_3E8}$ also reaches higher $\enucdot$ values at the low end of the temperature range ($< 0.5 \times 10^9 \mathrm{K}$). There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain reaction pathways that are not present in the cooler $\tt{F500}$ run. Note that, although $\tt{F500\_3E8}$ is the hottest run with a modified initial temperature configuration to form a distinct flame front, the highest rotation $\tt{F1000}$ flame actually reaches higher temperatures (see Figure~\ref{fig:maxima}, left panel) as well as higher $\bar{A}$ (see Figure~\ref{fig:time_series_1000}).

In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ 
is so hot that the organized flame structure is lost. The fuel layer undergoes stable burning and 
essentially simmers in place, rather than forming a flame. This process has been observed in 
neutron stars with accretion rates that exceed the Eddington limit 
\citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model we have essentially 
recreated the conditions found in such stars with high accretion rates. This model burns so 
strongly that it is only run out to $40~\mathrm{ms}$. After an initial period when the burning 
moves across the domain, residual burning continues and eventually ignites the entire fuel layer at 
late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three 
snapshots taken in the last ten seconds of the simulation. In Figure~\ref{fig:4e8_stacked_enuc}, 
$\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ 
across the domain and at heights up to $\sim 0.5\times 10^4~\mathrm{cm}$. There is still 
significant burning occurring even higher, with $\enucdot$ reaching $\sim 
10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 
10^4~\mathrm{cm}$. For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum 
$\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ 
(see Figure~\ref{fig:maxima}, right panel), even at the latest timesteps ($\sim 
70~\mathrm{ms}$, when the flame is most developed). Significant $\enucdot$ values for 
all other runs are confined to the flame front, rather than spanning the entire domain.

@@

 Though $\tt{F500\_4E8}$ does not form a distinct burning front, it does achieve greatly enhanced burning, as shown in the Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plot (right panel). Not only is $\tt{F500\_4E8}$ significantly hotter overall compared to all other runs, but there is a large causally connected region across a wide range of $\bar{A}$. This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures. The reaction pathways are also very disrupted for $\tt{F500\_4E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more vigorous mixing. Indeed, this appears to be the case looking at Figure~\ref{fig:4e8_stacked_abar} and the right panel of Figure~\ref{fig:uv_hot} (discussed further in section~\ref{ssec:temp_prop}). The disrupted reaction pathways resemble those found in Figure \ref{fig:abar} for the highest rotation $\tt{F1000}$ run, though they are even more dramatically disrupted for $\tt{F500\_4E8}$. $\tt{F500\_4E8}$ is also noticeably hotter than $\tt{F1000}$, even though it is run for significantly less time ($t = 40~\mathrm{ms}$ vs $t = 100~\mathrm{ms}$). Although the $\tt{F500\_4E8}$ run is clearly a special case in that it develops steady burning across the domain rather than a propagating flame, enhanced burning in this hottest run aligns with results from the other simulations with differing initial temperature structures. As $T_{\mathrm{star}}$ is increased, the overall behavior and propagation of the flame is significantly altered, implying that burning is very temperature-sensitive. We explore flame propagation further in Section~\ref{ssec:temp_prop}.

@@

The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K ($\tt{F500\_4e8}$; right panel) is significantly greater in magnitude than for the cooler $\tt{F500\_3e8}$ (middle panel) and $\tt{F500}$ (left panel) runs. The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the $\tt{F500\_4e8}$ flame's vertical $z$-velocity $v$ is also significantly larger. This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is two to three orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}. Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e. where $\dot{e}_{\rm nuc}$ is greatest) that can be clearly seen for the hotter runs at this time ($40~\mathrm{ms}$). As can be seen in Figure~\ref{fig:uv} (which was plotted at $100~\mathrm{ms}$), this does develop more at later times. Similar to what we saw when comparing the runs with different rotation rates, it would therefore appear that it takes longer for the flame to develop when $T_{\rm star}$ is cooler.

\begin{figure}[t]
    \centering
    \plotone{urho_plot_hotter}
    \caption{\label{fig:urho_hot}Phase plot of the horizontal $r$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$ (the latest time that $\tt{F500\_4E8}$ is run out to). The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
\end{figure}

>>>>

%======================================================================
% Abstract and Keywords
%======================================================================
\begin{abstract}
We continue to investigate laterally propagating flames in type I X-ray bursts using
fully compressible hydrodynamics simulations.  In the current study we
relax the approximations used previously and explore the effect of
rotation rate and the thermal structure of the neutron star on the
flame.  We find that at lower rotation rates the flame becomes harder to ignite,
whereas at higher rotation rates the nuclear burning is enhanced and the flame propagates steadily.
We also find that higher temperature atmospheres lead to faster
flames and observe acceleration of the flame front as it propagates
through the atmosphere.  All of the software used for these
simulations is freely available.

@@

We investigate the effect of rotation rate on the flame. With the
exception of IGR J17480-2446 (\citet{altamirano2010discovery}, spinning at $11~\mathrm{Hz}$), most observations of
XRBs come from sources with rotation rates of $200-600~\mathrm{Hz}$
\citep{bilous:2019,galloway:2020}. There are a number of
factors that could explain this lack of observations below $200~\mathrm{Hz}$. It
could be that there is some physical process which inhibits the flame
ignition and/or spread at lower rotation rates. It could be that burst
oscillations at lower rotation rates are smaller in amplitude and
therefore more difficult to detect. It could be that it does not have
anything to do with the flame at all, but that neutron stars in
accreting low mass X-ray binaries rarely have rotation rates below $200~\mathrm{Hz}$ 
\MarginPar{look into this?}.

@@

We use the Castro hydrodynamics code \citep{castro,castro_joss} and
the simulation framework introduced in \citet{flame_wave1}.  The
current simulations are all performed in a two-dimensional
axisymmetric geometry.  For these axisymmetric simulations, we add an
additional geometric source term from \citet{bernard-champmartin} that
captures the effects of the divergence of the flux operating on the
azimuthal unit vector.  This term is a small correction, but was
missing from our previous simulations.  The simulation framework
initializes a fuel layer in hydrostatic equilibrium, laterally
blending a hot model on the left side of the domain (the coordinate
origin) and a cool model on the right.  The initial temperature
gradient between the hot and cool fluids drives a laterally
propagating flame through the cool fuel.  In our original set of
calculations (in \citet{flame_wave1}), in order to make the 
simulations computationally feasible we artificially boosted the flame speed by adjusting the
conductivity and reaction rate to produce a flame moving 5--10$\times$
faster than the nominal laminar flame speed.  We also used high
rotation rates to reduce the lateral lengthscale at which the Coriolis
force balances the lateral flame spreading in order to reduce the size 
of the simulation domain. The port of \castro\ to GPUs~\citep{sc20_gpu} 
significantly improved its overall performance, enabling
us to run these new simulations without the previous approximations while
continuing to resolve the burning front. For these
simulations, we no longer boost the flame speed---the true
conductivities (taken from \citealt{Timmes00}) and reaction rates are
used. We are also able to use slower, more physically realistic rotation rates.
  We continue to use a 13-isotope $\alpha$-chain to describe the
helium burning.

@@

\begin{figure}
	\plotone{maxima_vs_time.pdf}
	\caption{\label{fig:maxima}Estimates of temperature (left panel) and nuclear energy generation 
	rate (right panel) in the burning region as functions of time. The quantities on the vertical 
	axes are the mass-weighted averages defined in Equation~\ref{eqn:top_percentile_avg}.}
\end{figure}

For some of our analysis, we would like to have a means of estimating the temperature ($T$) and nuclear 
energy generation rate ($\enucdot$) in the burning region of each simulation. For this purpose, we define 
the mass weighted average $\repavg{Q}$ of some quantity $Q$ to be
\begin{equation}
	\label{eqn:top_percentile_avg}
	\repavg{Q} \equiv \frac{\sum_{c_i} m(c_i) Q(c_i)}{\sum_{c_i} m(c_i)}; ~c_i \in C_{99}(Q).
\end{equation}
\noindent Here, $C_{99}(Q)$ is the set of grid cells with $Q$ values in the top percentile, and 
$Q(c_i)$ is the value of $Q$ in cell $c_i$\MarginPar{Please define what m(c i) here!}. By using 
$\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire simulation domain 
allows us to track changes over the domain as a whole rather than at a single localized 
point. This will therefore be a better reflection of the overall behavior of the flame rather than of a 
single localized fluctuation. Figure \ref{fig:maxima} shows $\repavg{T}$ and $\repavg{\enucdot}$ as functions of time for 
the subset of our runs that achieve a propagating flame. This figure is referenced throughout the 
subsequent sections.

@@

We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms 
of temperature but differing rotation rates. We saw in \citet{flame_wave1} that 
increasing the neutron star rotation rate reduces the horizontal lengthscale of the flame. An 
estimate of this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius 
may be thought of as the scale over which the balance between the Coriolis force and horizontal 
pressure gradient becomes important, and is approximately given by
\begin{equation}
	\label{eqn:rossby}
	L_R \approx \frac{\sqrt{g H_0}}{\Omega},
\end{equation}
where $g$ is the gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
neutron star rotation rate. In Figure \ref{fig:time_series_enuc_500} and Figure 
\ref{fig:time_series_enuc_1000}, we use the $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates. Taking the edge at greatest radius of the bright teal/green region where the most significant energy generation is occurring as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be 
reduced compared to the lower rotation $500~\mathrm{Hz}$ run ($\tt{F500}$). From Equation 
\ref{eqn:rossby}, we can see that increasing the rotation rate from $500~\mathrm{Hz}$ to 
$1000~\mathrm{Hz}$ should decrease $L_R$ by a factor of two, and that the greater confinement from 
the Coriolis force should reduce the horizontal extent of the flame by a similar factor. However, 
the Rossby radius is only an approximate measure of this horizontal lengthscale, and in our 
simulations we see that this scaling does not work so well for all rotation rates. The simulations 
seem to follow the theoretical scaling described in Equation \ref{eqn:rossby} more closely at 
higher rotation rates ($1000~\mathrm{Hz}$ and higher), based on the results of \citet{flame_wave1}. {\color{blue} Perhaps these new un-boosted runs have not had sufficient time to reach steady propagation, in contrast to the boosted runs in \citet{flame_wave1} that reached steady propagation more quickly.} \MarginPar{does the equation for the Rossby radius only hold for flames with steady propagation?}

@@

The {\tt F500} and {\tt F1000} runs both qualitatively resemble the flame structure in 
\citet{flame_wave1} --- a laterally propagating flame that is lifted off of the bottom of the fuel 
layer --- but they differ in their burning structures. Figures~\ref{fig:time_series_500} and 
\ref{fig:time_series_1000} show time series of the mean molecular weight, $\bar{A}$, for the {\tt 
F500} and {\tt F1000} runs.  Compared to those in \cite{flame_wave1}, ashes behind the flame do not 
reach as high atomic weights. This is not surprising, since those previous runs artificially 
boosted the reaction rates.  Comparing these two new runs, the burning is much more evolved for the 
higher rotation rate, and the ash is actually able to move ahead of the flame front (visible in the 
Figure  \ref{fig:time_series_1000} $100~\mathrm{ms}$ snapshot). We believe that this is because the increased 
rotation better confines the initial perturbation and subsequent expansion from the burning, 
increasing the temperature and density in the flame front such that the reaction rate increases, 
which allows the reactions to progress further. The $\dot{e}_\mathrm{nuc}$ plots in Figure 
\ref{fig:time_series_enuc_1000} also support this interpretation, with the region of the 
flame front nearest to the crust in the {\tt F1000} run reaching higher $\dot{e}_\mathrm{nuc}$ values 
than for the {\tt F500} run in Figure~\ref{fig:time_series_enuc_500}. In contrast to {\tt F500} 
and {\tt F1000}, the lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition 
for {\tt F250} also aligns with the reasoning given above, with the lower rotation in this case 
potentially leading to insufficient confinement such that the temperature and density required for 
ignition is not achieved. In this scenario, another source of confinement (e.g.\ magnetic fields, see \citep{art-2016-cavecchi-etal}) 
would need to take over at lower rotation rates to allow a burst to occur, at least for the initial 
model used here. {\color{blue} Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), it is also possible that we simply cannot confine the flame sufficiently with our current domain width.} We see in Figure~\ref{fig:flame_speeds_1} that the {\tt F500} flame took longer to achieve steady propagation than the {\tt F1000} flame. It may therefore also be that we did not run our simulation for long enough to see the {\tt F250} flame achieve the conditions required for ignition and steady propagation.

Burning in the {\tt F500} and {\tt F1000} runs is concentrated in a dense region with circular 
motion. In Figure~\ref{fig:urho}, which compares the horizontal $r$-velocity $u$, density $\rho$ and the 
nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the {\tt F250}, {\tt F500}, and {\tt F1000} 
runs, most of the burning for each of the simulations occurs in a high density region $\rho > 3 
\times 10^5 {\rm g\,cm^{-3}}$. The fluid in this dense, high energy generation region undergoes 
vortical motion, shown in the Figure~\ref{fig:uv} phase plots comparing $u$, the vertical $z$-velocity $v$ and $\dot{e}_{\rm 
nuc}$. This most likely corresponds to the leading edge of the flame where fresh fuel is being entrained. \MarginPar{not sure if we should keep this sentence about entrainment...}
This feature is not developed in the $250~\mathrm{Hz}$ flame in Figure~\ref{fig:uv} {\color{blue} (left panel)}; it could potentially develop 
at later times (past the point at which we terminated our simulation), or the flame could just 
fizzle out and fail to ignite entirely.

@@

Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures ($T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$) at $t = 70~\mathrm{ms}$. {\color{blue}We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front. $\tt{F500\_3E8}$ (Figure~\ref{fig:compare_500Hz_abar}, bottom panel) --- the hottest run to form a clear burning front ---} has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}). {\color{blue} It also reaches slightly higher $\bar{A}$ values than the two cooler runs. The Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plots of $\tt{F500}$ (left) and $\tt{F500\_3E8}$ (middle) also reflect these $\bar{A}$ features, with $\tt{F500\_3E8}$ reaching slightly higher $\bar{A}$ values. $\tt{F500\_3E8}$ also reaches higher $\enucdot$ values at the low end of the temperature range ($< 0.5 \times 10^9 \mathrm{K}$). There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain reaction pathways that are not present in the cooler $\tt{F500}$ run. Note that, although $\tt{F500\_3E8}$ is the hottest run with a modified initial temperature configuration to form a distinct flame front, the highest rotation $\tt{F1000}$ flame actually reaches higher temperatures (see Figure~\ref{fig:maxima}, left panel) as well as higher $\bar{A}$ (see Figure~\ref{fig:time_series_1000}).}

In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ 
is so hot that the organized flame structure is lost. The fuel layer undergoes stable burning and 
essentially simmers in place, rather than forming a flame. This process has been observed in 
neutron stars with accretion rates that exceed the Eddington limit 
\citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model we have essentially 
recreated the conditions found in such stars with high accretion rates. This model burns so 
strongly that it is only run out to $40~\mathrm{ms}$. After an initial period when the burning 
moves across the domain, residual burning continues and eventually ignites the entire fuel layer at 
late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three 
snapshots taken in the last ten seconds of the simulation. In Figure~\ref{fig:4e8_stacked_enuc}, 
$\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ 
across the domain and at heights up to $\sim 0.5\times 10^4~\mathrm{cm}$. There is still 
significant burning occurring even higher, with $\enucdot$ reaching $\sim 
10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 
10^4~\mathrm{cm}$. For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum 
$\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ 
{\color{blue} (see Figure~\ref{fig:maxima}, right panel)}, even at the latest timesteps ($\sim 
70~\mathrm{ms}$, when the flame is most developed). {\color{blue} Significant $\enucdot$ values for 
all other runs are confined to the flame front, rather than spanning the entire domain}.

@@

 {\color{blue} Though $\tt{F500\_4E8}$ does not form a distinct burning front, it does achieve greatly enhanced burning, as shown in the Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plot (right panel). Not only is $\tt{F500\_4E8}$ significantly hotter overall compared to all other runs}, but there is a large causally connected region across a wide range of $\bar{A}$. This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures. The reaction pathways are also very disrupted for $\tt{F500\_4E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more vigorous mixing. {\color{blue}Indeed, this appears to be the case looking at Figure~\ref{fig:4e8_stacked_abar} and the right panel of Figure~\ref{fig:uv_hot} (discussed further in section~\ref{ssec:temp_prop}). The disrupted reaction pathways resemble those found in Figure \ref{fig:abar} for the highest rotation $\tt{F1000}$ run, though they are even more dramatically disrupted for $\tt{F500\_4E8}$. $\tt{F500\_4E8}$ is also noticeably hotter than $\tt{F1000}$, even though it is run for significantly less time ($t = 40~\mathrm{ms}$ vs $t = 100~\mathrm{ms}$). Although the $\tt{F500\_4E8}$ run is clearly a special case in that it develops steady burning across the domain rather than a propagating flame, enhanced burning in this hottest run aligns with results from the other simulations with differing initial temperature structures. As $T_{\mathrm{star}}$ is increased, the overall behavior and propagation of the flame is significantly altered, implying that burning is very temperature-sensitive. We explore flame propagation further in Section~\ref{ssec:temp_prop}.}

@@

The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K {\color{blue} ($\tt{F500\_4e8}$; right panel)} is significantly greater in magnitude than for the cooler {\color{blue}$\tt{F500\_3e8}$ (middle panel) and $\tt{F500}$ (left panel)} runs. The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the {\color{blue}$\tt{F500\_4e8}$ flame's vertical $z$-velocity $v$} is also significantly larger. This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is two to three orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}. Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e. where $\dot{e}_{\rm nuc}$ is greatest) that can be clearly seen for the hotter runs at this time ($40~\mathrm{ms}$). As can be seen in Figure~\ref{fig:uv} (which was plotted at $100~\mathrm{ms}$), this does develop more at later times. Similar to what we saw when comparing the runs with different rotation rates, it would therefore appear that it takes longer for the flame to develop when $T_{\rm star}$ is cooler.

\begin{figure}[t]
    \centering
    \plotone{urho_plot_hotter}
    \caption{\label{fig:urho_hot}Phase plot of the horizontal $r$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$ {\color{blue} (the latest time that $\tt{F500\_4E8}$ is run out to). The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
\end{figure}

\begin{figure}[t]
    \centering
    \plotone{uv_plot_hotter}
    \caption{\label{fig:uv_hot}Phase plot of the horizontal $r$-velocity $u$, vertical $z$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$. {\color{blue} The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
\end{figure}
